---
title: "Renewable Energy Infrastructure Overlap Report"
output: pdf_document
params:
  study_name: NA
  track_results: NA
  track_points: NA
  wind_pts: NA
  solar_polys: NA
  buffer_distance_m: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(sf)
library(patchwork)

res <- params$track_results
n_tracks_total <- if (!is.null(res) && nrow(res) > 0) length(unique(res$track_id)) else 0
```

```{r table-helpers}
# Plain formatted number, comma-separated if requested, NA -> "-". Used by
# the highlights section and any spot that needs a formatted value outside
# of a table (tables themselves use knitr::kable's own digits/format.args,
# not this).
plain_num <- function(x, digits = 0, big_mark = "") {
  ifelse(is.na(x), "-",
         formatC(round(x, digits), format = "f", digits = digits, big.mark = big_mark))
}

# Plain "Yes"/"No"/"-" for a logical column.
yn <- function(x) ifelse(is.na(x), "-", ifelse(x, "Yes", "No"))

# Human-readable version of a raw Movebank field name (e.g.
# "height_above_ground_level" -> "height above ground level").
pretty_field_name <- function(x) ifelse(is.na(x), "-", gsub("_", " ", x))

na_dash <- function(x) ifelse(is.na(x), "-", x)
```

## Study

**Study name:** `r ifelse(is.na(params$study_name), "(not available)", params$study_name)`

**Overlap buffer used:** `r params$buffer_distance_m` m

**Number of tracks assessed:** `r nrow(res)`

**Taxa present:** `r if (all(is.na(res$taxon))) "(not available)" else paste(sort(unique(na.omit(res$taxon))), collapse = ", ")`

## Data quality summary

```{r quality-table}
quality_display <- res %>%
  mutate(
    pct_days_monitored = pmin(100, 100 * n_days_monitored / duration_days)
  ) %>%
  transmute(
    `Track ID` = track_id,
    `Taxon` = na_dash(taxon),
    `N locations` = n_locations,
    `Duration (days)` = duration_days,
    `Days monitored` = n_days_monitored,
    `Days monitored (%)` = pct_days_monitored,
    `Median fix interval (h)` = median_fix_interval_hours
  )

knitr::kable(
  quality_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 1, 0, 0, 1)
)
```

*"Days monitored" is the count of distinct calendar days represented by the
point set used for the infrastructure overlap test below (the same set
`enable_thinning`, if on, would subsample); "Days monitored (%)" is that
count as a percentage of "Duration (days)," so a track sampled sparsely or
with large data gaps relative to its own start/end dates will show up
lower here. This App does not apply a fixed pass/fail quality
threshold -- what counts as sufficient data for a follow-up impact
analysis depends on the taxon and research question. Use these metrics,
together with the exposure metrics below, to make that judgment for your
own study.*

## Key exposure highlights

```{r highlights, results='asis'}
if (nrow(res) > 0) {
  pct_wind_overlap  <- 100 * mean(res$overlaps_wind, na.rm = TRUE)
  pct_solar_overlap <- 100 * mean(res$overlaps_solar, na.rm = TRUE)
  n_wind_overlap    <- sum(res$overlaps_wind, na.rm = TRUE)
  n_solar_overlap   <- sum(res$overlaps_solar, na.rm = TRUE)

  cat(sprintf(
    "- **%.0f%%** of tracks (%d of %d) came within the buffer of at least one **wind turbine** at some point (point-confirmed or interpolated path).\n",
    pct_wind_overlap, n_wind_overlap, nrow(res)
  ))
  cat(sprintf(
    "- **%.0f%%** of tracks (%d of %d) came within the buffer of at least one **solar array** at some point.\n",
    pct_solar_overlap, n_solar_overlap, nrow(res)
  ))

  top_wind_idx <- which.max(res$pct_wind_points_in_buffer)
  if (length(top_wind_idx) > 0 && !is.na(res$pct_wind_points_in_buffer[top_wind_idx]) &&
      res$pct_wind_points_in_buffer[top_wind_idx] > 0) {
    cat(sprintf(
      "- Track **%s** spent the most time near **wind** infrastructure: **%.0f%%** of its tested fixes fell within the buffer, across **%s** separate visit(s), longest lasting **%.1f h**.\n",
      res$track_id[top_wind_idx], res$pct_wind_points_in_buffer[top_wind_idx],
      plain_num(res$n_wind_visit_bouts[top_wind_idx], 0),
      ifelse(is.na(res$wind_longest_buffer_bout_hours[top_wind_idx]), 0,
             res$wind_longest_buffer_bout_hours[top_wind_idx])
    ))
  }

  top_solar_idx <- which.max(res$pct_solar_points_in_buffer)
  if (length(top_solar_idx) > 0 && !is.na(res$pct_solar_points_in_buffer[top_solar_idx]) &&
      res$pct_solar_points_in_buffer[top_solar_idx] > 0) {
    cat(sprintf(
      "- Track **%s** spent the most time near **solar** infrastructure: **%.0f%%** of its tested fixes fell within the buffer, across **%s** separate visit(s), longest lasting **%.1f h**.\n",
      res$track_id[top_solar_idx], res$pct_solar_points_in_buffer[top_solar_idx],
      plain_num(res$n_solar_visit_bouts[top_solar_idx], 0),
      ifelse(is.na(res$solar_longest_buffer_bout_hours[top_solar_idx]), 0,
             res$solar_longest_buffer_bout_hours[top_solar_idx])
    ))
  }

  if (n_wind_overlap == 0 && n_solar_overlap == 0) {
    cat("- No tracks in this study showed any overlap with the checked infrastructure.\n")
  }
} else {
  cat("No tracks were assessed in this run.")
}
```

```{r exposure-bar-chart, fig.width=7, fig.height=max(3, 0.35 * n_tracks_total + 1)}
exposure_long <- dplyr::bind_rows(
  res %>% transmute(track_id, Infrastructure = "Wind", pct = pct_wind_points_in_buffer),
  res %>% transmute(track_id, Infrastructure = "Solar", pct = pct_solar_points_in_buffer)
) %>%
  filter(!is.na(pct))

if (nrow(exposure_long) > 0) {
  ggplot(exposure_long, aes(x = reorder(track_id, pct), y = pct, fill = Infrastructure)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    coord_flip() +
    scale_fill_manual(values = c(Wind = "#4C72B0", Solar = "#DD8452")) +
    labs(x = NULL, y = "% of tested fixes within buffer",
         title = "Exposure intensity by track") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No exposure-intensity data available for this chart") +
    theme_void()
}
```

## Infrastructure overlap summary

Overlap evidence comes in two kinds: **point-confirmed** (a recorded GPS
fix actually fell within the buffer -- direct evidence) and **track-only**
(the interpolated path between two consecutive fixes passed within the
buffer, but neither fix itself did -- inferred, not confirmed, and most
relevant when fix intervals are coarse). "N turbines"/"N arrays" below
combines both kinds (any facility with either type of evidence); the
exposure-intensity columns (points/days/visits/distance) are
**point-confirmed only**, since a "visit" or "day spent nearby" should
reflect an actual recorded detection, not just an inferred path crossing.

### Wind

```{r overlap-table-wind, results='asis'}
wind_display <- res %>%
  transmute(
    `Track ID` = track_id,
    `Overlap?` = yn(overlaps_wind),
    `N turbines` = n_turbines_overlap,
    `Pts in buffer (%)` = pct_wind_points_in_buffer,
    `Days in buffer (%)` = pct_wind_days_in_buffer,
    `N visits` = n_wind_visit_bouts,
    `Longest visit (h)` = wind_longest_buffer_bout_hours,
    `Nearest (m)` = nearest_turbine_dist_m,
    `Median dist (m)` = wind_dist_median_m,
    `Timing` = na_dash(wind_temporal_relation)
  )

# \small/\normalsize are base LaTeX font-size commands (no package needed),
# used here (instead of kableExtra's scale_down) just to give this
# 10-column table a bit more room on the page.
cat("\\small\n\n")
print(knitr::kable(
  wind_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
))
cat("\n\n\\normalsize\n\n")
```

*"Pts/Days in buffer (%)" are percentages of the point set tested for
overlap (see the quality-summary note above for what that set is); "Median
dist" is the median distance to the nearest turbine across ALL fixes (not
just the single closest approach shown in "Nearest"), reflecting how close
the track ranges *typically*, not only at its closest point. Full detail
(including mean distance and buffer-specific fix interval) is in the CSV
artifact.*

```{r height-table, results='asis'}
height_used <- "wind_height_field_used" %in% names(res) && any(!is.na(res$wind_height_field_used))

if (height_used) {
  cat("**Height relative to turbines** (points within the horizontal buffer only -- see README's Height analysis section for the full methodology and its limits)\n\n")

  height_display <- res %>%
    mutate(pct_rotor_zone = ifelse(n_wind_points_in_buffer > 0,
                                    100 * n_wind_points_in_rotor_zone / n_wind_points_in_buffer,
                                    NA_real_)) %>%
    transmute(
      `Track ID` = track_id,
      `Height field used` = pretty_field_name(wind_height_field_used),
      `True AGL?` = yn(wind_height_is_agl),
      `N in rotor zone` = n_wind_points_in_rotor_zone,
      `In rotor zone (%)` = pct_rotor_zone,
      `Min height (m)` = wind_height_min_m,
      `Median height (m)` = wind_height_median_m,
      `Max height (m)` = wind_height_max_m
    )

  cat("\\small\n\n")
  print(knitr::kable(
    height_display, booktabs = TRUE,
    digits = c(0, 0, 0, 0, 0, 0, 0, 0)
  ))
  cat("\n\n\\normalsize\n\n")

  cat("\n\n*\"In rotor zone (%)\" is the percentage of buffer-zone fixes falling within a nearby turbine's rotor-swept height range, and (like \"N in rotor zone\") is only ever populated when \"True AGL?\" is Yes -- otherwise height is reported for context only, since ellipsoid/mean-sea-level height cannot be safely compared to a ground-relative hazard zone without a ground-elevation correction this App does not perform.*\n")
}
```

### Solar

```{r overlap-table-solar, results='asis'}
solar_display <- res %>%
  transmute(
    `Track ID` = track_id,
    `Overlap?` = yn(overlaps_solar),
    `N arrays` = n_solar_arrays_overlap,
    `Pts in buffer (%)` = pct_solar_points_in_buffer,
    `Days in buffer (%)` = pct_solar_days_in_buffer,
    `N visits` = n_solar_visit_bouts,
    `Longest visit (h)` = solar_longest_buffer_bout_hours,
    `Nearest (m)` = nearest_solar_dist_m,
    `Median dist (m)` = solar_dist_median_m,
    `Timing` = na_dash(solar_temporal_relation)
  )

cat("\\small\n\n")
print(knitr::kable(
  solar_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
))
cat("\n\n\\normalsize\n\n")
```

*Timing categories: "pre-operational" (tracking entirely before the facility's
operational year), "post-operational" (entirely after), "spanning" (before
and after), "mixed" (track overlaps multiple facilities with different
operational years).*

## Map

```{r map-setup}
# wind_pts/solar_polys arrive in whatever the App's working CRS was
# (Conus Albers when solar was checked, otherwise a local UTM zone) --
# reproject everything to WGS84 once, here, for plotting.
wind_wgs84 <- if (!is.null(params$wind_pts) && nrow(params$wind_pts) > 0) {
  sf::st_transform(params$wind_pts, 4326)
} else NULL

solar_wgs84 <- if (!is.null(params$solar_polys) && nrow(params$solar_polys) > 0) {
  sf::st_transform(params$solar_polys, 4326)
} else NULL

tracks_wgs84 <- if (!is.null(params$track_points) && nrow(params$track_points) > 0) {
  params$track_points
} else NULL

# Build one LINESTRING per track from tracks_wgs84's points, relying on
# RFunction.R having already sorted them chronologically before thinning --
# a scattered cloud of same-colored points is very hard to read as a route
# once a dense infrastructure layer (e.g. thousands of turbines) is drawn
# underneath it; a connected line reads immediately as "this is where the
# animal went," even against a busy backdrop. Tracks with fewer than 2
# points can't form a line and are simply skipped here (they still show up
# via the point layer, which is always drawn in addition to this one).
build_track_lines <- function(pts_sf) {
  if (is.null(pts_sf) || nrow(pts_sf) == 0) return(NULL)
  ids <- unique(pts_sf$track_id)
  lines <- lapply(ids, function(tid) {
    sub <- pts_sf[pts_sf$track_id == tid, ]
    if (nrow(sub) < 2) return(NULL)
    coords <- sf::st_coordinates(sub)[, c("X", "Y"), drop = FALSE]
    sf::st_sfc(sf::st_linestring(coords), crs = sf::st_crs(sub))
  })
  keep <- !vapply(lines, is.null, logical(1))
  if (!any(keep)) return(NULL)
  sf::st_sf(track_id = ids[keep], geometry = do.call(c, lines[keep]))
}

tracks_lines_wgs84 <- build_track_lines(tracks_wgs84)

# A handful of evenly-spaced tick breaks (rather than each panel's default,
# which for a narrow-extent track can pack in enough decimal-degree labels
# to visually overlap) -- applied to every map below via
# scale_x/y_continuous(), which coord_sf() still respects.
few_breaks <- function(n = 3) scales::breaks_pretty(n = n)

# Crop a layer to a bbox-derived extent, tolerating empty results/edge cases
# rather than letting a cosmetic map crop fail the whole report render.
safe_crop <- function(x, extent) {
  if (is.null(x) || nrow(x) == 0) return(NULL)
  tryCatch(sf::st_crop(x, extent), error = function(e) NULL)
}

# Build a bbox-derived extent (WGS84) around a set of points, with a margin
# so points aren't plotted flush against the figure edge. Falls back to a
# fixed minimum margin for single-point or very tight tracks.
bbox_extent <- function(pts, margin_frac = 0.15, min_margin_deg = 0.1) {
  bbox <- sf::st_bbox(pts)
  margin_x <- max(min_margin_deg, diff(bbox[c("xmin", "xmax")]) * margin_frac)
  margin_y <- max(min_margin_deg, diff(bbox[c("ymin", "ymax")]) * margin_frac)
  sf::st_as_sfc(bbox + c(-margin_x, -margin_y, margin_x, margin_y), crs = 4326)
}

# Crop the infrastructure layers to the study's actual extent for the
# overview map -- wind_pts/solar_polys as received here can cover a much
# larger area than any single study (solar_polys in particular is always
# the full nationwide GM-SEUS extract, not pre-cropped to the study).
overview_extent <- if (!is.null(tracks_wgs84)) bbox_extent(tracks_wgs84) else NULL
wind_overview  <- if (!is.null(overview_extent)) safe_crop(wind_wgs84, overview_extent) else wind_wgs84
solar_overview <- if (!is.null(overview_extent)) safe_crop(solar_wgs84, overview_extent) else solar_wgs84

# Per-track maps are paginated (see map-per-track below) into fixed-size
# groups rather than crammed into one giant image -- a single grid sized to
# fit however many tracks the study has would just get scaled down to fit
# the page regardless of how tall it's asked to be, so panels in a
# large study (e.g. 50+ individuals) would shrink to illegible thumbnails.
# Splitting into fixed-size pages keeps panel size legible and consistent
# no matter how many tracks the study has -- more tracks just means more
# pages, not smaller panels.
tracks_per_page <- 12
ncol_per_page <- 3
```

```{r map-overview, fig.width=7, fig.height=6}
p <- ggplot()
has_layer <- FALSE

if (!is.null(solar_overview) && nrow(solar_overview) > 0) {
  p <- p + geom_sf(data = solar_overview, fill = "#DD8452", color = NA, alpha = 0.5)
  has_layer <- TRUE
}
if (!is.null(wind_overview) && nrow(wind_overview) > 0) {
  # Turbine markers are deliberately tiny and semi-transparent -- a dense
  # wind corridor can mean tens of thousands of points in view at once
  # (e.g. a wide-ranging migratory study), and at normal point size/opacity
  # they blot out everything else on the map, tracks included. Small +
  # translucent still reads clearly as "lots of turbines here" without
  # drowning out the animal's actual path.
  p <- p + geom_sf(data = wind_overview, color = "#4C72B0", shape = 17,
                    size = 0.3, alpha = 0.35, stroke = 0)
  has_layer <- TRUE
}
if (!is.null(tracks_lines_wgs84) && nrow(tracks_lines_wgs84) > 0) {
  p <- p + geom_sf(data = tracks_lines_wgs84, aes(color = track_id), linewidth = 0.6)
  has_layer <- TRUE
}
if (!is.null(tracks_wgs84) && nrow(tracks_wgs84) > 0) {
  p <- p + geom_sf(data = tracks_wgs84, aes(color = track_id), size = 0.4, alpha = 0.8) +
    scale_color_viridis_d(option = "turbo")
  has_layer <- TRUE
}

if (has_layer) {
  p + theme_minimal(base_size = 11) +
    scale_x_continuous(breaks = few_breaks(4)) +
    scale_y_continuous(breaks = few_breaks(4)) +
    labs(title = "Tracks with nearby wind turbines (triangles) and solar arrays (polygons)",
         color = "Track ID") +
    theme(legend.position = "bottom")
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No data available for map") +
    theme_void()
}
```

*If tracks are geographically dispersed (e.g. separate breeding/wintering
areas, or multiple distinct individuals/locations), this overview will span
that full extent -- see the per-track maps below for a zoomed view of each
individual's own local area.*

### Individual track maps

Each panel below is independently zoomed to its own track's extent,
condensed into a single grid (via `patchwork`) rather than one page per
individual -- this keeps the report a manageable length for studies with
many tracked individuals (e.g. large migratory-bird studies) while still
giving each track its own local view. (Note: `facet_wrap(scales = "free")`
can't be used here -- ggplot2's `coord_sf()`, which every `geom_sf()` plot
uses, doesn't support per-panel free scales within a single `facet_wrap()`
call, so each panel below is built as its own fully independent plot with
its own zoom, then composed into a grid, rather than one shared faceted
plot.) The animal's path is drawn as a connected black line rather than
scattered points, so it stays legible even over a dense nearby
infrastructure layer; wind turbines in particular are deliberately drawn
small and translucent, since a busy wind-development corridor can put
hundreds of them in a single panel. Panels are paginated into fixed-size
groups (a consistent number of panels per page) rather than one giant grid
sized to fit the whole study -- otherwise a large study (e.g. 50+
individuals) would still get squeezed onto a single page and every panel
would shrink to an illegible thumbnail; pagination keeps panel size
legible and consistent regardless of how many tracks the study has, at the
cost of more pages for larger studies.

```{r map-per-track, fig.width=9, fig.height=1 + 3.2 * ceiling(tracks_per_page / ncol_per_page), results='asis'}
if (!is.null(tracks_wgs84) && nrow(tracks_wgs84) > 0) {
  track_ids_facet <- unique(tracks_wgs84$track_id)

  # One fully independent ggplot per track (own coord_sf, own zoom extent)
  # -- see the note above on why this can't be a single faceted plot.
  build_one_track_plot <- function(tid) {
    track_i <- tracks_wgs84[tracks_wgs84$track_id == tid, ]
    line_i  <- if (!is.null(tracks_lines_wgs84)) {
      tracks_lines_wgs84[tracks_lines_wgs84$track_id == tid, ]
    } else NULL
    extent_i <- bbox_extent(track_i, margin_frac = 0.25, min_margin_deg = 0.05)

    solar_i <- safe_crop(solar_wgs84, extent_i)
    wind_i  <- safe_crop(wind_wgs84, extent_i)

    taxon_i <- res$taxon[res$track_id == tid][1]
    title_i <- if (!is.na(taxon_i) && nzchar(taxon_i)) paste0(tid, " (", taxon_i, ")") else tid

    p_i <- ggplot()
    if (!is.null(solar_i) && nrow(solar_i) > 0) {
      p_i <- p_i + geom_sf(data = solar_i, fill = "#DD8452", color = NA, alpha = 0.5)
    }
    if (!is.null(wind_i) && nrow(wind_i) > 0) {
      # Small + translucent, same reasoning as the overview map -- a
      # zoomed-in panel over a dense wind corridor can still have hundreds
      # of nearby turbines, which at full size/opacity would bury the
      # track underneath.
      p_i <- p_i + geom_sf(data = wind_i, color = "#4C72B0", shape = 17,
                            size = 0.5, alpha = 0.4, stroke = 0)
    }
    if (!is.null(line_i) && nrow(line_i) > 0) {
      p_i <- p_i + geom_sf(data = line_i, color = "black", linewidth = 0.6)
    }

    p_i +
      geom_sf(data = track_i, color = "black", size = 0.4, alpha = 0.8) +
      scale_x_continuous(breaks = few_breaks(3)) +
      scale_y_continuous(breaks = few_breaks(3)) +
      labs(title = title_i) +
      theme_minimal(base_size = 9) +
      theme(axis.text = element_text(size = 6),
            axis.text.x = element_text(angle = 45, hjust = 1),
            plot.title = element_text(size = 9))
  }

  track_plots <- lapply(track_ids_facet, build_one_track_plot)

  # Paginate: tracks_per_page panels per page (see map-setup for why), each
  # page printed as its own figure with a \newpage (base LaTeX, no package
  # needed) between pages -- results='asis' is required for that raw
  # command to pass through rather than being escaped, and for print() to
  # actually emit each page's plot inside this loop (only a chunk's final
  # unassigned expression auto-prints otherwise).
  n_total <- length(track_plots)
  page_starts <- seq(1, n_total, by = tracks_per_page)

  for (start in page_starts) {
    end <- min(start + tracks_per_page - 1, n_total)
    print(patchwork::wrap_plots(track_plots[start:end], ncol = ncol_per_page))
    if (end < n_total) cat("\n\n\\newpage\n\n")
  }
} else {
  cat("No track location data available.")
}
```

## Data sources

- **Wind:** U.S. Wind Turbine Database (USWTDB), USGS/LBNL/American Clean
  Power Association. CC BY 3.0.
- **Solar:** Ground-Mounted Solar Energy in the United States (GM-SEUS),
  Stid et al., Michigan State University. CC BY 4.0 (v2.0, Zenodo record
  19581821).
