---
title: "Renewable Energy Infrastructure Overlap Report"
output: pdf_document
params:
  study_name: NA
  track_results: NA
  track_points: NA
  wind_pts: NA
  solar_polys: NA
  buffer_distance_m: NA
  height_points: NA
  min_locations: NA
  min_duration_days: NA
  max_fix_interval_hours: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(sf)
library(patchwork)

res <- params$track_results
n_tracks_total <- if (!is.null(res) && nrow(res) > 0) length(unique(res$track_id)) else 0

# Dynamic figure height so the quality scorecard (one row per track, faceted
# by metric) stays readable as track count grows, without wasting space for
# small studies -- same reasoning as the exposure chart's height below.
fig_h_scorecard <- max(3, 0.35 * n_tracks_total + 1.5)
```

```{r table-helpers}
# Plain formatted number, comma-separated if requested, NA -> "-". Used by
# the highlights section and any spot that needs a formatted value outside
# of a table (tables themselves use knitr::kable's own digits/format.args,
# not this).
plain_num <- function(x, digits = 0, big_mark = "") {
  ifelse(is.na(x), "-",
         formatC(round(x, digits), format = "f", digits = digits, big.mark = big_mark))
}

# Plain "Yes"/"No"/"-" for a logical column.
yn <- function(x) ifelse(is.na(x), "-", ifelse(x, "Yes", "No"))

# Human-readable version of a raw Movebank field name (e.g.
# "height_above_ground_level" -> "height above ground level").
pretty_field_name <- function(x) ifelse(is.na(x), "-", gsub("_", " ", x))

na_dash <- function(x) ifelse(is.na(x), "-", x)
```

## Study

**Study name:** `r ifelse(is.na(params$study_name), "(not available)", params$study_name)`

**Overlap buffer used:** `r params$buffer_distance_m` m

**Number of tracks assessed:** `r nrow(res)`

**Taxa present:** `r if (all(is.na(res$taxon))) "(not available)" else paste(sort(unique(na.omit(res$taxon))), collapse = ", ")`

## Data quality summary

```{r quality-scorecard, fig.width=9, fig.height=fig_h_scorecard}
# MoveApps may pass an explicit NULL (not just omit the argument) for an
# unset optional numeric setting -- as.numeric(NULL) is a zero-length
# vector, not NA, which would silently break data.frame()'s row recycling
# below (a length-0 column against length-n_tracks_total ones). Collapse
# NULL/zero-length/non-numeric to a proper scalar NA instead.
safe_num <- function(x) {
  v <- suppressWarnings(as.numeric(x))
  if (length(v) == 0 || is.na(v[1])) NA_real_ else v[1]
}

scorecard_data <- dplyr::bind_rows(
  data.frame(track_id = res$track_id, metric = "N locations",
             value = res$n_locations, threshold = safe_num(params$min_locations)),
  data.frame(track_id = res$track_id, metric = "Duration (days)",
             value = res$duration_days, threshold = safe_num(params$min_duration_days)),
  data.frame(track_id = res$track_id, metric = "Median fix interval (h)",
             value = res$median_fix_interval_hours, threshold = safe_num(params$max_fix_interval_hours))
)
scorecard_data$metric <- factor(scorecard_data$metric,
                                 levels = c("N locations", "Duration (days)", "Median fix interval (h)"))

ggplot(scorecard_data, aes(x = value, y = track_id)) +
  geom_vline(aes(xintercept = threshold), linetype = "dashed", color = "grey40") +
  geom_point(size = 2.8, color = "#4C72B0") +
  facet_wrap(~metric, scales = "free_x", nrow = 1) +
  labs(x = NULL, y = "Track ID") +
  theme_minimal()
```

```{r quality-table}
quality_display <- res %>%
  mutate(
    pct_days_monitored = pmin(100, 100 * n_days_monitored / duration_days)
  ) %>%
  transmute(
    `Track ID` = track_id,
    `Taxon` = na_dash(taxon),
    `N locations` = n_locations,
    `Duration (days)` = duration_days,
    `Days monitored` = n_days_monitored,
    `Days monitored (%)` = pct_days_monitored,
    `Median fix interval (h)` = median_fix_interval_hours
  )

knitr::kable(
  quality_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 1, 0, 0, 1)
)
```

*"Days monitored" is the count of distinct calendar days represented by the
point set used for the infrastructure overlap test below; "Days monitored (%)" 
is that count as a percentage of "Duration (days)."

## Key exposure highlights

```{r highlights, results='asis'}
if (nrow(res) > 0) {
  pct_wind_overlap  <- 100 * mean(res$overlaps_wind, na.rm = TRUE)
  pct_solar_overlap <- 100 * mean(res$overlaps_solar, na.rm = TRUE)
  n_wind_overlap    <- sum(res$overlaps_wind, na.rm = TRUE)
  n_solar_overlap   <- sum(res$overlaps_solar, na.rm = TRUE)

  cat(sprintf(
    "- **%.0f%%** of tracks (%d of %d) came within the buffer of at least one **wind turbine** at some point (point-confirmed or interpolated path).\n",
    pct_wind_overlap, n_wind_overlap, nrow(res)
  ))
  cat(sprintf(
    "- **%.0f%%** of tracks (%d of %d) came within the buffer of at least one **solar array** at some point.\n",
    pct_solar_overlap, n_solar_overlap, nrow(res)
  ))

  top_wind_idx <- which.max(res$pct_wind_points_in_buffer)
  if (length(top_wind_idx) > 0 && !is.na(res$pct_wind_points_in_buffer[top_wind_idx]) &&
      res$pct_wind_points_in_buffer[top_wind_idx] > 0) {
    cat(sprintf(
      "- Track **%s** spent the most time near **wind** infrastructure: **%.0f%%** of its tested fixes fell within the buffer, across **%s** separate visit(s), longest lasting **%.1f h**.\n",
      res$track_id[top_wind_idx], res$pct_wind_points_in_buffer[top_wind_idx],
      plain_num(res$n_wind_visit_bouts[top_wind_idx], 0),
      ifelse(is.na(res$wind_longest_buffer_bout_hours[top_wind_idx]), 0,
             res$wind_longest_buffer_bout_hours[top_wind_idx])
    ))
  }

  top_solar_idx <- which.max(res$pct_solar_points_in_buffer)
  if (length(top_solar_idx) > 0 && !is.na(res$pct_solar_points_in_buffer[top_solar_idx]) &&
      res$pct_solar_points_in_buffer[top_solar_idx] > 0) {
    cat(sprintf(
      "- Track **%s** spent the most time near **solar** infrastructure: **%.0f%%** of its tested fixes fell within the buffer, across **%s** separate visit(s), longest lasting **%.1f h**.\n",
      res$track_id[top_solar_idx], res$pct_solar_points_in_buffer[top_solar_idx],
      plain_num(res$n_solar_visit_bouts[top_solar_idx], 0),
      ifelse(is.na(res$solar_longest_buffer_bout_hours[top_solar_idx]), 0,
             res$solar_longest_buffer_bout_hours[top_solar_idx])
    ))
  }

  if (n_wind_overlap == 0 && n_solar_overlap == 0) {
    cat("- No tracks in this study showed any overlap with the checked infrastructure.\n")
  }
} else {
  cat("No tracks were assessed in this run.")
}
```

```{r exposure-bar-chart, fig.width=7, fig.height=max(3, 0.35 * n_tracks_total + 1)}
exposure_long <- dplyr::bind_rows(
  res %>% transmute(track_id, Infrastructure = "Wind", pct = pct_wind_points_in_buffer),
  res %>% transmute(track_id, Infrastructure = "Solar", pct = pct_solar_points_in_buffer)
) %>%
  filter(!is.na(pct))

if (nrow(exposure_long) > 0) {
  # Fixed 0-100 scale (rather than one that zooms to the data range) so a
  # bar's length is always comparable to the full range of possible
  # exposure, and a reader skimming the chart shape alone -- without
  # reading the axis text -- can't mistake a genuinely tiny percentage for
  # a large one. Since a fixed 0-100 scale can make already-small values
  # nearly invisible as bars, each bar also gets its own text label with
  # the exact value, so the number is still legible even when the bar
  # itself is a sliver.
  ggplot(exposure_long, aes(x = reorder(track_id, pct), y = pct, fill = Infrastructure)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6) +
    geom_text(
      aes(label = sprintf("%.1f%%", pct)),
      position = position_dodge(width = 0.7),
      hjust = -0.15, size = 3, color = "gray30"
    ) +
    coord_flip(clip = "off") +
    scale_y_continuous(limits = c(0, 100), expand = expansion(mult = c(0, 0.08))) +
    scale_fill_manual(values = c(Wind = "#4C72B0", Solar = "#DD8452")) +
    labs(x = NULL, y = "% of tested fixes within buffer",
         title = "Exposure intensity by track") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom",
          plot.margin = margin(5.5, 22, 5.5, 5.5))
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No exposure-intensity data available for this chart") +
    theme_void()
}
```

## Infrastructure overlap summary

Overlap evidence is either **1) point-confirmed**, where a recorded GPS
fix actually fell within the buffer, or **2) track-only**, where the 
interpolated path between two consecutive fixes passed within the
buffer, but neither did. "N turbines"/"N arrays" below
combines both kinds, while the exposure-intensity columns 
(points/days/visits/distance) are **point-confirmed only**.

### Overview

```{r temporal-breakdown-setup}
temporal_data <- dplyr::bind_rows(
  if (any(!is.na(res$wind_temporal_relation))) {
    as.data.frame(table(relation = res$wind_temporal_relation)) %>%
      dplyr::mutate(infra = "Wind") %>%
      dplyr::rename(n = Freq)
  },
  if (any(!is.na(res$solar_temporal_relation))) {
    as.data.frame(table(relation = res$solar_temporal_relation)) %>%
      dplyr::mutate(infra = "Solar") %>%
      dplyr::rename(n = Freq)
  }
)
has_temporal_data <- !is.null(temporal_data) && nrow(temporal_data) > 0
```

```{r temporal-breakdown, fig.width=6, fig.height=4, eval=has_temporal_data}
ggplot(temporal_data, aes(x = infra, y = n, fill = relation)) +
  geom_col(position = "stack", width = 0.5) +
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3.5) +
  labs(x = NULL, y = "Number of tracks", fill = "Temporal relation") +
  theme_minimal()
```

```{r temporal-breakdown-note, results='asis'}
if (!has_temporal_data) {
  cat("*No temporal relation data available (no overlapping infrastructure",
      "with a known operational year).*\n")
}
```

*Timing categories: "pre-operational" (tracking entirely before the facility's
operational year), "post-operational" (entirely after), "spanning" (before
and after), "mixed" (track overlaps multiple facilities with different
operational years).*

### Wind

```{r overlap-table-wind, results='asis'}
wind_display <- res %>%
  transmute(
    `Track ID` = track_id,
    `Overlap?` = yn(overlaps_wind),
    `N turbines` = n_turbines_overlap,
    `Pts in buffer (%)` = pct_wind_points_in_buffer,
    `Days in buffer (%)` = pct_wind_days_in_buffer,
    `N visits` = n_wind_visit_bouts,
    `Longest visit (h)` = wind_longest_buffer_bout_hours,
    `Nearest (m)` = nearest_turbine_dist_m,
    `Median dist (m)` = wind_dist_median_m,
    `Timing` = na_dash(wind_temporal_relation)
  )

# \small/\normalsize are base LaTeX font-size commands (no package needed),
# used here (instead of kableExtra's scale_down) just to give this
# 10-column table a bit more room on the page.
cat("\\small\n\n")
print(knitr::kable(
  wind_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
))
cat("\n\n\\normalsize\n\n")
```

```{r height-table, results='asis'}
height_used <- "wind_height_field_used" %in% names(res) && any(!is.na(res$wind_height_field_used))

if (height_used) {
  cat("**Height relative to turbines** (points within the horizontal buffer only -- see README's Height analysis section for the full methodology and its limits)\n\n")

  height_display <- res %>%
    mutate(pct_rotor_zone = ifelse(n_wind_points_in_buffer > 0,
                                    100 * n_wind_points_in_rotor_zone / n_wind_points_in_buffer,
                                    NA_real_)) %>%
    transmute(
      `Track ID` = track_id,
      `Height field used` = pretty_field_name(wind_height_field_used),
      `True AGL?` = yn(wind_height_is_agl),
      `N in rotor zone` = n_wind_points_in_rotor_zone,
      `In rotor zone (%)` = pct_rotor_zone,
      `Min height (m)` = wind_height_min_m,
      `Median height (m)` = wind_height_median_m,
      `Max height (m)` = wind_height_max_m
    )

  cat("\\small\n\n")
  print(knitr::kable(
    height_display, booktabs = TRUE,
    digits = c(0, 0, 0, 0, 0, 0, 0, 0)
  ))
  cat("\n\n\\normalsize\n\n")

  cat("\n\n*\"In rotor zone (%)\" is the percentage of buffer-zone fixes falling within a nearby turbine's rotor-swept height range, and (like \"N in rotor zone\") is only ever populated when \"True AGL?\" is Yes -- otherwise height is reported for context only, since ellipsoid/mean-sea-level height cannot be safely compared to a ground-relative hazard zone without a ground-elevation correction this App does not perform.*\n")
}
```

```{r height-points-setup}
# Per-point height/distance detail (one row per tested fix with a non-NA
# height), shared by both the wind and solar height figures below. NULL/
# zero-row/missing-column all collapse to the same "nothing to plot" state,
# so downstream chunks only need to check height_pts_available.
hp <- params$height_points
height_pts_available <- !is.null(hp) && is.data.frame(hp) && nrow(hp) > 0 &&
  "height_m" %in% names(hp)

# Wind figure scope: buffer-zone fixes only, matching the height-table above
# ("points within the horizontal buffer only").
wind_height_pts <- if (height_pts_available) {
  hp %>% filter(wind_in_buffer, !is.na(height_m))
} else {
  tibble::tibble()
}
n_wind_facet_tracks <- max(1, length(unique(wind_height_pts$track_id)))

# Solar figure scope: ALL tested fixes with a known distance to the nearest
# array, not just buffer-zone ones -- the point of this figure is to show
# how height varies with proximity, including well outside the buffer, so
# restricting to buffer-zone fixes would cut off exactly the comparison
# that makes it interesting.
solar_height_pts <- if (height_pts_available) {
  hp %>% filter(!is.na(height_m), !is.na(solar_dist_m))
} else {
  tibble::tibble()
}
```

```{r height-dist-wind, fig.width=8, fig.height=max(3, 1.9 * ceiling(n_wind_facet_tracks / 3) + 1)}
if (height_pts_available && nrow(wind_height_pts) > 0) {
  wind_height_pts_plot <- wind_height_pts %>%
    mutate(rotor_zone = dplyr::case_when(
      is.na(wind_in_rotor_zone) ~ "Unclassified",
      wind_in_rotor_zone ~ "In rotor zone",
      TRUE ~ "Outside rotor zone"
    ))

  ggplot(wind_height_pts_plot, aes(x = height_m, fill = rotor_zone)) +
    geom_histogram(bins = 30, color = "white", linewidth = 0.1) +
    facet_wrap(~track_id, scales = "free", ncol = 3) +
    scale_fill_manual(
      values = c("In rotor zone" = "#C44E52", "Outside rotor zone" = "#4C72B0",
                 "Unclassified" = "gray70"),
      breaks = c("In rotor zone", "Outside rotor zone", "Unclassified")
    ) +
    labs(x = "Height (m)", y = "N fixes", fill = NULL,
         title = "Height distribution near wind turbines, by track") +
    theme_minimal(base_size = 10) +
    theme(legend.position = "bottom", strip.text = element_text(size = 8))
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No per-point height data available for this chart") +
    theme_void()
}
```

### Solar

```{r overlap-table-solar, results='asis'}
solar_display <- res %>%
  transmute(
    `Track ID` = track_id,
    `Overlap?` = yn(overlaps_solar),
    `N arrays` = n_solar_arrays_overlap,
    `Pts in buffer (%)` = pct_solar_points_in_buffer,
    `Days in buffer (%)` = pct_solar_days_in_buffer,
    `N visits` = n_solar_visit_bouts,
    `Longest visit (h)` = solar_longest_buffer_bout_hours,
    `Nearest (m)` = nearest_solar_dist_m,
    `Median dist (m)` = solar_dist_median_m,
    `Timing` = na_dash(solar_temporal_relation)
  )

cat("\\small\n\n")
print(knitr::kable(
  solar_display, booktabs = TRUE, format.args = list(big.mark = ","),
  digits = c(0, 0, 0, 0, 0, 0, 1, 0, 0, 0)
))
cat("\n\n\\normalsize\n\n")
```

```{r height-table-solar, results='asis'}
solar_height_used <- "solar_height_field_used" %in% names(res) && any(!is.na(res$solar_height_field_used))

if (solar_height_used) {
  cat("**Height relative to panels** (points within the horizontal buffer only -- see README's Height analysis section for the full methodology and its limits). Unlike wind's rotor-swept zone, ground-mounted solar panels aren't a rotating strike hazard, so this is not a pass/fail hazard classification -- it simply reports whether animals near an array were flying close to panel level or well above it, against a single assumed panel height (GM-SEUS provides no per-array height field; see the App's README for the assumed value).\n\n")

  height_display_solar <- res %>%
    mutate(pct_at_or_below_panel = ifelse(n_solar_points_in_buffer > 0,
                                           100 * n_solar_points_at_or_below_panel_height / n_solar_points_in_buffer,
                                           NA_real_)) %>%
    transmute(
      `Track ID` = track_id,
      `Height field used` = pretty_field_name(solar_height_field_used),
      `True AGL?` = yn(solar_height_is_agl),
      `At/below panel ht (%)` = pct_at_or_below_panel,
      `N above panel ht` = n_solar_points_above_panel_height,
      `Median height (m)` = solar_height_median_m,
      `Median vs. panel (m)` = solar_height_median_relative_to_panel_m
    )

  cat("\\small\n\n")
  print(knitr::kable(
    height_display_solar, booktabs = TRUE,
    digits = c(0, 0, 0, 0, 0, 0, 1)
  ))
  cat("\n\n\\normalsize\n\n")

  assumed_ht <- res$solar_assumed_panel_height_m[!is.na(res$solar_assumed_panel_height_m)][1]
  if (!is.na(assumed_ht)) {
    cat(sprintf("\n\n*Assumed panel height: **%.0f m**. \"At/below panel ht (%%)\" is the percentage of buffer-zone fixes at or below that height; \"Median vs. panel (m)\" is the median fix height minus the assumed panel height (positive = flying above the panels, negative = at/below). Both are only ever populated when \"True AGL?\" is Yes -- otherwise height is reported for context only, since ellipsoid/mean-sea-level height cannot be safely compared to panel height without a ground-elevation correction this App does not perform.*\n", assumed_ht))
  } else {
    cat("\n\n*No fixes had a true-AGL height, so no assumed-panel-height comparison could be made -- see the field/AGL columns above for context only.*\n")
  }
}
```

```{r height-dist-solar, fig.width=7, fig.height=5}
if (height_pts_available && nrow(solar_height_pts) > 0) {
  # Cap the x-axis for readability -- GM-SEUS is a nationwide layer, so
  # st_nearest_feature() always finds *some* array, occasionally hundreds of
  # km away for a track nowhere near solar development. coord_cartesian()
  # only clips the displayed range; it doesn't drop data from the smooth.
  dist_cap <- max(
    params$buffer_distance_m * 1.5,
    stats::quantile(solar_height_pts$solar_dist_m, 0.95, na.rm = TRUE),
    na.rm = TRUE
  )

  ggplot(solar_height_pts, aes(x = solar_dist_m, y = height_m)) +
    # Small + translucent, same reasoning as the turbine markers on the map
    # -- a long-running or unthinned track can mean many thousands of
    # points, which at full size/opacity would just be a solid blob.
    geom_point(size = 0.5, alpha = 0.25, color = "#DD8452") +
    geom_smooth(method = "loess", se = TRUE, color = "gray20", linewidth = 0.6) +
    geom_vline(xintercept = params$buffer_distance_m, linetype = "dashed", color = "gray40") +
    annotate("text", x = params$buffer_distance_m, y = Inf, label = "buffer distance",
             angle = 90, vjust = -0.5, hjust = 1.1, size = 3, color = "gray40") +
    coord_cartesian(xlim = c(0, dist_cap)) +
    labs(x = "Distance to nearest solar array (m)", y = "Height (m)",
         title = "Flight height vs. distance to nearest solar array (all tracks pooled)") +
    theme_minimal(base_size = 11)
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No per-point height/distance data available for this chart") +
    theme_void()
}
```

## Map

```{r map-setup}
# wind_pts/solar_polys arrive in whatever the App's working CRS was
# (Conus Albers when solar was checked, otherwise a local UTM zone) --
# reproject everything to WGS84 once, here, for plotting.
wind_wgs84 <- if (!is.null(params$wind_pts) && nrow(params$wind_pts) > 0) {
  sf::st_transform(params$wind_pts, 4326)
} else NULL

solar_wgs84 <- if (!is.null(params$solar_polys) && nrow(params$solar_polys) > 0) {
  sf::st_transform(params$solar_polys, 4326)
} else NULL

tracks_wgs84 <- if (!is.null(params$track_points) && nrow(params$track_points) > 0) {
  params$track_points
} else NULL

# Build one LINESTRING per track from tracks_wgs84's points, relying on
# RFunction.R having already sorted them chronologically before thinning --
# a scattered cloud of same-colored points is very hard to read as a route
# once a dense infrastructure layer (e.g. thousands of turbines) is drawn
# underneath it; a connected line reads immediately as "this is where the
# animal went," even against a busy backdrop. Tracks with fewer than 2
# points can't form a line and are simply skipped here (they still show up
# via the point layer, which is always drawn in addition to this one).
build_track_lines <- function(pts_sf) {
  if (is.null(pts_sf) || nrow(pts_sf) == 0) return(NULL)
  ids <- unique(pts_sf$track_id)
  lines <- lapply(ids, function(tid) {
    sub <- pts_sf[pts_sf$track_id == tid, ]
    if (nrow(sub) < 2) return(NULL)
    coords <- sf::st_coordinates(sub)[, c("X", "Y"), drop = FALSE]
    sf::st_sfc(sf::st_linestring(coords), crs = sf::st_crs(sub))
  })
  keep <- !vapply(lines, is.null, logical(1))
  if (!any(keep)) return(NULL)
  sf::st_sf(track_id = ids[keep], geometry = do.call(c, lines[keep]))
}

tracks_lines_wgs84 <- build_track_lines(tracks_wgs84)

# A handful of evenly-spaced tick breaks (rather than each panel's default,
# which for a narrow-extent track can pack in enough decimal-degree labels
# to visually overlap) -- applied to every map below via
# scale_x/y_continuous(), which coord_sf() still respects.
few_breaks <- function(n = 3) scales::breaks_pretty(n = n)

# Crop a layer to a bbox-derived extent, tolerating empty results/edge cases
# rather than letting a cosmetic map crop fail the whole report render.
#
# Solar arrays reach this function already reprojected to WGS84 (lon/lat).
# sf's default spherical (S2) geometry engine enforces ring-orientation
# rules that are stricter than -- and different from -- the planar OGC
# validity already checked/repaired in fetch_gmseus() (which runs in the
# projected working CRS). A polygon that is perfectly valid there can still
# trip an S2 "invalid orientation"-style error once it's in geographic
# coordinates, which st_crop() hits internally via st_intersection(). That
# error was previously caught and silently turned into NULL here -- which
# is exactly why solar arrays were disappearing from the map even though
# the overlap counts (computed earlier, in the projected CRS, before this
# function ever runs) were correct. Since cropping to a bounding box for a
# static figure is a purely planar operation anyway, S2 is switched off for
# the duration of the crop (and always restored after, even on error).
safe_crop <- function(x, extent) {
  if (is.null(x) || nrow(x) == 0) return(NULL)
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)
  tryCatch(sf::st_crop(x, extent), error = function(e) {
    warning(paste("safe_crop: dropping a layer after crop error:", conditionMessage(e)))
    NULL
  })
}

# A fixed-size marker at each solar array's centroid, layered on top of the
# polygon fill. A geom_sf() point's size is a fixed physical size on the
# page regardless of the map's zoom level, whereas a several-hundred-meter
# array's polygon fill is not -- across a whole study's extent (the
# overview map) that fill can render as a sub-pixel sliver, effectively
# invisible, even directly under a fix (see the "Nearest (m) = 0" rows in
# the Solar table). The marker stays visible exactly when the fill isn't.
# At per-track (already zoomed-in) scale, where the polygon is plenty
# visible on its own, the same small fixed size keeps this a subtle accent
# rather than a second competing layer -- deliberately not scaled up there.
#
# st_point_on_surface() (not st_centroid()) guarantees a point that
# actually lies on the polygon: an array clipped at a crop's edge can end
# up an irregular or split multipolygon shape, for which a true geometric
# centroid can fall outside the shape entirely (e.g. a crescent-shaped
# remainder after clipping). Wrapped with S2 off for the same reason as
# safe_crop() above -- this runs on the same post-crop WGS84 geometries.
safe_centroid <- function(x) {
  if (is.null(x) || nrow(x) == 0) return(NULL)
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)
  tryCatch(suppressWarnings(sf::st_point_on_surface(x)), error = function(e) {
    warning(paste("safe_centroid: dropping centroid markers after error:", conditionMessage(e)))
    NULL
  })
}

# Build a bbox-derived extent (WGS84) around a set of points, with a margin
# so points aren't plotted flush against the figure edge. Falls back to a
# fixed minimum margin for single-point or very tight tracks.
bbox_extent <- function(pts, margin_frac = 0.15, min_margin_deg = 0.1) {
  bbox <- sf::st_bbox(pts)
  margin_x <- max(min_margin_deg, diff(bbox[c("xmin", "xmax")]) * margin_frac)
  margin_y <- max(min_margin_deg, diff(bbox[c("ymin", "ymax")]) * margin_frac)
  sf::st_as_sfc(bbox + c(-margin_x, -margin_y, margin_x, margin_y), crs = 4326)
}

# Crop the infrastructure layers to the study's actual extent for the
# overview map -- wind_pts/solar_polys as received here can cover a much
# larger area than any single study (solar_polys in particular is always
# the full nationwide GM-SEUS extract, not pre-cropped to the study).
overview_extent <- if (!is.null(tracks_wgs84)) bbox_extent(tracks_wgs84) else NULL
wind_overview  <- if (!is.null(overview_extent)) safe_crop(wind_wgs84, overview_extent) else wind_wgs84
solar_overview <- if (!is.null(overview_extent)) safe_crop(solar_wgs84, overview_extent) else solar_wgs84

# Per-track maps are paginated (see map-per-track below) into fixed-size
# groups rather than crammed into one giant image -- a single grid sized to
# fit however many tracks the study has would just get scaled down to fit
# the page regardless of how tall it's asked to be, so panels in a
# large study (e.g. 50+ individuals) would shrink to illegible thumbnails.
# Splitting into fixed-size pages keeps panel size legible and consistent
# no matter how many tracks the study has -- more tracks just means more
# pages, not smaller panels.
tracks_per_page <- 12
ncol_per_page <- 3
```

```{r map-overview, fig.width=7, fig.height=6}
p <- ggplot()
has_layer <- FALSE

if (!is.null(solar_overview) && nrow(solar_overview) > 0) {
  p <- p + geom_sf(data = solar_overview, fill = "#DD8452", color = NA, alpha = 0.5)
  solar_overview_centroids <- safe_centroid(solar_overview)
  if (!is.null(solar_overview_centroids) && nrow(solar_overview_centroids) > 0) {
    p <- p + geom_sf(data = solar_overview_centroids, shape = 18, size = 1,
                      color = "#DD8452", alpha = 0.85)
  }
  has_layer <- TRUE
}
if (!is.null(wind_overview) && nrow(wind_overview) > 0) {
  # Turbine markers are deliberately tiny and semi-transparent -- a dense
  # wind corridor can mean tens of thousands of points in view at once
  # (e.g. a wide-ranging migratory study), and at normal point size/opacity
  # they blot out everything else on the map, tracks included. Small +
  # translucent still reads clearly as "lots of turbines here" without
  # drowning out the animal's actual path.
  p <- p + geom_sf(data = wind_overview, color = "#4C72B0", shape = 17,
                    size = 0.3, alpha = 0.35, stroke = 0)
  has_layer <- TRUE
}
if (!is.null(tracks_lines_wgs84) && nrow(tracks_lines_wgs84) > 0) {
  p <- p + geom_sf(data = tracks_lines_wgs84, aes(color = track_id), linewidth = 0.6)
  has_layer <- TRUE
}
if (!is.null(tracks_wgs84) && nrow(tracks_wgs84) > 0) {
  p <- p + geom_sf(data = tracks_wgs84, aes(color = track_id), size = 0.4, alpha = 0.8) +
    scale_color_viridis_d(option = "turbo")
  has_layer <- TRUE
}

if (has_layer) {
  p + theme_minimal(base_size = 11) +
    scale_x_continuous(breaks = few_breaks(4)) +
    scale_y_continuous(breaks = few_breaks(4)) +
    labs(color = "Track ID") +
    theme(legend.position = "bottom")
} else {
  ggplot() +
    annotate("text", x = 0, y = 0, label = "No data available for map") +
    theme_void()
}
```

*Wind turbines are shown as small triangles and solar arrays as shaded
polygons with a diamond marker at their centroids. If tracks are geographically 
dispersed (e.g. separate breeding/wintering areas, or multiple distinct 
individuals/locations), thisoverview will span that full extent. See the 
per-track maps below for a zoomed view of each individual's own local area.*

### Individual track maps

```{r map-per-track, fig.width=9, fig.height=1 + 3.2 * ceiling(tracks_per_page / ncol_per_page), results='asis'}
if (!is.null(tracks_wgs84) && nrow(tracks_wgs84) > 0) {
  track_ids_facet <- unique(tracks_wgs84$track_id)

  # One fully independent ggplot per track (own coord_sf, own zoom extent)
  # -- see the note above on why this can't be a single faceted plot.
  build_one_track_plot <- function(tid) {
    track_i <- tracks_wgs84[tracks_wgs84$track_id == tid, ]
    line_i  <- if (!is.null(tracks_lines_wgs84)) {
      tracks_lines_wgs84[tracks_lines_wgs84$track_id == tid, ]
    } else NULL
    extent_i <- bbox_extent(track_i, margin_frac = 0.25, min_margin_deg = 0.05)

    solar_i <- safe_crop(solar_wgs84, extent_i)
    wind_i  <- safe_crop(wind_wgs84, extent_i)

    taxon_i <- res$taxon[res$track_id == tid][1]
    title_i <- if (!is.na(taxon_i) && nzchar(taxon_i)) paste0(tid, " (", taxon_i, ")") else tid

    p_i <- ggplot()
    if (!is.null(solar_i) && nrow(solar_i) > 0) {
      p_i <- p_i + geom_sf(data = solar_i, fill = "#DD8452", color = NA, alpha = 0.5)
      solar_i_centroids <- safe_centroid(solar_i)
      if (!is.null(solar_i_centroids) && nrow(solar_i_centroids) > 0) {
        # Slightly smaller than the overview map's marker (0.8 vs 1) -- these
        # panels are physically smaller on the page (multiple per row), so
        # the same absolute size would read as more dominant here.
        p_i <- p_i + geom_sf(data = solar_i_centroids, shape = 18, size = 0.8,
                              color = "#DD8452", alpha = 0.8)
      }
    }
    if (!is.null(wind_i) && nrow(wind_i) > 0) {
      # Small + translucent, same reasoning as the overview map -- a
      # zoomed-in panel over a dense wind corridor can still have hundreds
      # of nearby turbines, which at full size/opacity would bury the
      # track underneath.
      p_i <- p_i + geom_sf(data = wind_i, color = "#4C72B0", shape = 17,
                            size = 0.5, alpha = 0.4, stroke = 0)
    }
    if (!is.null(line_i) && nrow(line_i) > 0) {
      p_i <- p_i + geom_sf(data = line_i, color = "black", linewidth = 0.6)
    }

    p_i +
      geom_sf(data = track_i, color = "black", size = 0.4, alpha = 0.8) +
      scale_x_continuous(breaks = few_breaks(3)) +
      scale_y_continuous(breaks = few_breaks(3)) +
      labs(title = title_i) +
      theme_minimal(base_size = 9) +
      theme(axis.text = element_text(size = 6),
            axis.text.x = element_text(angle = 45, hjust = 1),
            plot.title = element_text(size = 9))
  }

  track_plots <- lapply(track_ids_facet, build_one_track_plot)

  # Paginate: tracks_per_page panels per page (see map-setup for why), each
  # page printed as its own figure with a \newpage (base LaTeX, no package
  # needed) between pages -- results='asis' is required for that raw
  # command to pass through rather than being escaped, and for print() to
  # actually emit each page's plot inside this loop (only a chunk's final
  # unassigned expression auto-prints otherwise).
  n_total <- length(track_plots)
  page_starts <- seq(1, n_total, by = tracks_per_page)

  for (start in page_starts) {
    end <- min(start + tracks_per_page - 1, n_total)
    print(patchwork::wrap_plots(track_plots[start:end], ncol = ncol_per_page))
    if (end < n_total) cat("\n\n\\newpage\n\n")
  }
} else {
  cat("No track location data available.")
}
```

## Data sources

- **Wind:** U.S. Wind Turbine Database (USWTDB), USGS/LBNL/American Clean
  Power Association. CC BY 3.0.
- **Solar:** Ground-Mounted Solar Energy in the United States (GM-SEUS),
  Stid et al., Michigan State University. CC BY 4.0 (v2.0, Zenodo record
  19581821).